\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.

### 2012-04-23 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
CHANGES IN VERSION 1.5.3 -------------------------  o extractSpectra is now defunct <2012-04-20 Fri>  o caching full header in level 1; this is required when and MSnExp    instance with *many* spectra (created from many raw files) is    quantified - calling header(object) is a too big overhead compared    to actual reporter quantification. <2012-04-20 Fri>  o The header() method now uses the cached dataframe if level >= 1;    the (unexported) .header function can be used to generate the    dataframe using the assayData slot data. <2012-04-20 Fri>  o Setting processingData in MSnSet initialisat

The \Rfunction{plot2d} method takes one \Robject{MSnExp} instance as firs argument 
to produce retention time \textit{vs.} precursor MZ scatter plots. Points represent 
individual MS2 spectra and can be coloured based on 
precursor charge (with second argument \Robject{z="charge"}), 
total ion current (\Robject{z="tic"}),
number of peaks in the MS2 spectra \Robject{z="peaks.count"}) or, 
when multiple data files were loaded, file \Robject{z="file"}), 
as illustrated on figure \ref{fig:plot2d}. 
The lower right panel is produced for only a subset of proteins. 
See the method documentation for more details.

@@

The \Rfunction{plotDensity} method illustrates the distribution 
of several parameters of interest (see figure \ref{fig:plotDensity}).
Similarly to \Rfunction{plot2d}, the first argument is an \Robject{MSnExp} instance. 
The second is one of \Robject{precursor.mz}, \Robject{peaks.count} or \Robject{tic}, 
whose density will be plotted. 
An optional third argument specifies whether the x axes should be logged.

@@

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity threshold below which peaks are removed 
is defined by the \Robject{t} parameter. \Robject{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest non null intensity in any spectrum. 
We observe the effect of the \Rfunction{removePeaks} method by comparing 
total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{tic} method
before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
spectrum \texttt{X55}.

<<removePeaks,echo=TRUE,cache=FALSE>>=
experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
## total ion current
tic(itraqdata[["X55"]])
tic(experiment[["X55"]])
@

@@

\begin{figure}[p]
<<norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=7>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[6],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt,ttl="Non-normalised data")
.plot(qnt.max)
.plot(qnt.sum)
.plot(qnt.quant)
.plot(qnt.qrob)
.plot(qnt.vsn)
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

>>>>

The \Rfunction{plot2d} method takes one \Robject{MSnExp} instance as firs argument 
to produce retention time \textit{vs.} precursor MZ scatter plots. Points represent 
individual MS2 spectra and can be coloured based on 
precursor charge (with second argument \Robject{z="charge"}), 
total ion count (\Robject{z="ionCount"}),
number of peaks in the MS2 spectra \Robject{z="peaks.count"}) or, 
when multiple data files were loaded, file \Robject{z="file"}), 
as illustrated on figure \ref{fig:plot2d}. 
The lower right panel is produced for only a subset of proteins. 
See the method documentation for more details.

@@

The \Rfunction{plotDensity} method illustrates the distribution 
of several parameters of interest (see figure \ref{fig:plotDensity}).
Similarly to \Rfunction{plot2d}, the first argument is an \Robject{MSnExp} instance. 
The second is one of \Robject{precursor.mz}, \Robject{peaks.count} or \Robject{ionCount}, 
whose density will be plotted. 
An optional third argument specifies whether the x axes should be logged.

@@

There are several methods implemented to perform basic data manipulation. 
Low intensity peaks can be set to 0 with the \Rfunction{removePeaks} method 
from spectra or whole experiments. The intensity threshold below which peaks are removed 
is defined by the \Robject{t} parameter. \Robject{t} can be specified directly 
as a numeric. The default value is the character \texttt{"min"}, that will remove 
all peaks equal to the lowest non null intensity in any spectrum. 
We observe the effect of the \Rfunction{removePeaks} method by comparing 
total ion count (i.e. the total intensity in a spectrum) with the \Rfunction{ionCount} 
method before (object \Robject{itraqdata}) and after (object \Robject{experiment}) for 
spectrum \texttt{X55}.

<<removePeaks,echo=TRUE,cache=FALSE>>=
experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
## total ion current
ionCount(itraqdata[["X55"]])
ionCount(experiment[["X55"]])
@

@@

\begin{figure}[p]
<<norm-plot,fig=T,echo=F,tikz=T,external=F,width=5,height=7>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt,ttl="Non-normalised data")
.plot(qnt.max)
.plot(qnt.sum)
.plot(qnt.quant)
.plot(qnt.qrob)
.plot(qnt.vsn)
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}
